### 2023-09-26 doc/_experiment.qmd
small changes to experiment section

To determine the relative fitness of the ensemble of genotypes, a series of
growth-dilution cycles are performed on either a single or multiple
environments. In other words, the cultures are grown for some time; then, an
aliquot is inoculated into fresh media for the next growth cycle. This process
is repeated for roughly 4-7 cycles, depending on the initial abundances of the
mutants and their relative growth rates. The DNA barcodes are sequenced at the
end of each growth cycle to quantify the relative abundance of each of the
barcodes. We point the reader to @kinsler2020 for specific details on these
assays for *S. cerevisiae* and to @ascensao2023 for equivalent assays for *E.
coli*. @fig-01(B) presents a typical barcode trajectory where the darker barcode
trajectories represent the so-called *neutral lineages* that are genetically
equivalent to the untagged ancestor strain that initially dominates the culture.
These neutral lineages simplify the inference problem since the fitness of all
relevant barcodes is quantified with respect to these barcodes.

![**Typical competitive fitness experiment**. (A) Schematic of the typical
experimental design to determine the competitive fitness of an ensemble of
barcoded genotypes. Genotypes are pooled together and grown over multiple
growth-dilution cycles. At the end of each cycle, a sample is processed to
generate a library for amplicon sequencing. (B) Typical barcode trajectory
dataset. From each time point, the relative frequency of each barcode is
determined from the total number of reads. Different shades of blue represent
different relative fitness. Darker gray lines define the typical trajectory of
neutral lineages.](./figs/fig01){#fig-01}

>>>>

To determine the relative fitness of the ensemble of genotypes, a series of
growth-dilution cycles are performed on either a single or multiple
environments. In other words, the cultures are grown for some time; then, an
aliquot is inoculated into fresh media for the next growth cycle. This process
is repeated for roughly 4-7 cycles, depending on the initial abundances of the
mutants and their relative growth rates. The DNA barcodes are sequenced at the
end of each growth cycle to quantify the relative abundance of each of the
barcodes. We point the reader to @kinsler2020 for specific details on these
assays for *S. cerevisiae* and to @ascensao2023 for equivalent assays for *E.
coli*. @fig-01(B) presents a typical barcode trajectory where the darker barcode
trajectories represent the so-called *neutral lineages*, genetically equivalent
to the untagged ancestor strain that initially dominates the culture. These
neutral lineages simplify the inference problem since the fitness of all
relevant barcodes is quantified with respect to these barcodes.

![**Typical competitive fitness experiment**. (A) Schematic of the typical
experimental design to determine the competitive fitness of an ensemble of
barcoded genotypes. Genotypes are pooled together and grown over multiple
growth-dilution cycles. At the end of each cycle, a sample is processed to
generate a library for amplicon sequencing. (B) Typical barcode trajectory
dataset. From each time point, the relative frequency of each barcode is
determined from the total number of reads. Shades of blue represent different
relative fitness. Darker gray lines define the typical trajectory of neutral
lineages.](./figs/fig01){#fig-01}
